---
title: "CSDE 502- Homework 4"
author: "Carolyn Fan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
mailto: "cfan5@uw.edu"
output: 
  bookdown::html_document2:
    toc: TRUE
    number_sections: TRUE
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(magrittr)
library(knitr)
library(readstata13)
library(pander)
library(captioner)
library(Rmisc)
```


```{r, warning=FALSE, message=FALSE, echo=FALSE}
gradpain <- c(rep(0,11), rep(1,1), rep(2,2), rep(3,6), rep(4,8), rep(5,10), rep(6,8), rep(7,30), rep(8,10), rep(9,6), rep(10,2))
```

## Question 1: How many graduate students are in the sample? Use R code to determine this.
```{r, warning=FALSE, message=FALSE}
length(gradpain)

```
There are `r length(gradpain)` students in the sample. 

## Question 2: What is the sample mean?
```{r, warning=FALSE, message=FALSE}
mean(gradpain)
```
The sample mean of pain on a scale from 0 to 10 is `r mean(gradpain)`.

## Box 1
```{r, warning=FALSE, message=FALSE}
fx_box1 <- function (x){
  d.vec <- x
  n <- length(x)
  return(list("N"=n, "Mean"=mean(x)))
}

fx_box1(gradpain)

```

## Box 2
```{r, warning=FALSE, message=FALSE}
s <- sample(x=gradpain, size=100, replace=TRUE)

F <- NULL

set.seed(7)
for(i in seq(from=1, to=100, by=1)){
  fx_box1(gradpain)
  s <- sample(x=gradpain, size=100, replace=TRUE)
  m <- mean(s)
  F <- c(F, m)
}

mean(F)
sd(F)
```

## Question 3: What is the mean of these 100 means?
The mean is `r mean(F)`.

## Question 4: What is the standard deviation of these 100 means?
The standard deviation is `r sd(F)`.

## Box 3
```{r, warning=FALSE, message=FALSE}

D <- NULL

fx_box3 <-function(x, d.vec, n, num.samples){
  d.vec <- x
  n <- length(x)
  num.samples <- 100
  m <- mean(x)
  return(c(n, m))
}

for(i in seq(from=1, to=100, by=1)){
  fx_box3(gradpain)
  s <- sample(x=gradpain, size=100, replace=TRUE)
  num.samples <- 100
  m <- mean(s)
  D <- c(D, m)
  m2 <- mean(D)
  sd <- sd(D)
  ci <- CI(x=D, ci=0.95)
}
options(digits=3)

```

## Question 5: What does your function return for the mean of means?
The mean of means is `r m2`

## Question 6: What does your function return for the standard deviation of means?
The standard deviation is `r sd`

## Question 7: What does your function return for the 95% confidence interval around the mean?
The 95% CI, with the mean in the middle, is (`r ci`).


## Source Code
[cfan5_HW_4.rmd](cfan5_HW_4.Rmd)

```{r, comment='', echo=FALSE}
cat(readLines("cfan5_HW_4.rmd"), sep='\n')
```

